---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# scpopulation

## An R-package containing population data for the state of South Carolina.

<!-- badges: start -->
<!-- badges: end -->

The `scpopulation` is an open-source R-package published on GitHub. The package stores county level population data and projections from the South Carolina Office of Revenue and Fiscal Affairs (SC RFA). This website documents the scpopulation R-package and provides access to the data for the general public without requiring the use of R software.

This package was developed and is maintained by Dr. Charles Alex Pellett, while working at the South Carolina Department of Natural Resources (SC DNR) and subsequently the South Carolina Department of Environmental Services (SC DES). This information is for research and long-term water planning. None of the above guarantees the accuracy, reliability, or completeness of this information. The SC RFA, SC DNR, SC DES, and Dr. Pellett are NOT liable for any loss or damage arising from the use of this information.

As an R-package, `scpopulation` can be installed in an R terminal with the following code:

```{r eval=F}
# install.packages('devtools') 
devtools::install_github('https://github.com/capellett/scpopulation')
```

## Statewide population projection summary


### Multiple editions


### Extended projections


### Multiple scenarios


## County-level cohort-component projections

### Population pyramids
```{r}
### Show statewide plot with different projections over time.

## Show animated cohort-component plots.

## Show extended projections for several counties.

## Show extended statewide projections

## Provide searchable tables for download.


```
