---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# scpopulation

<!-- badges: start -->
<!-- badges: end -->

The scpopulation package stores county level population data and projections from the South Carolina Office of Revenue and Fiscal Affairs (SC RFA). The package was developed and is maintained by Dr. Charles Alex Pellett, while working at the South Carolina Department of Natural Resources (SC DNR) and subsequently the South Carolina Department of Environmental Services (SC DES). This information is for research and long-term water planning. None of the above guarantees the accuracy, reliability, or completeness of this information. The SC RFA, SC DNR, SC DES, and Dr. Pellett are NOT liable for any loss or damage arising from the use of this information.

```{r}
### Show statewide plot with different projections over time.

## Show animated cohort-component plots.

## Show extended projections for several counties.

## Show extended statewide projections

## Provide searchable tables for download.


```
